---
title: "Marketing_Project - Team 4"
output:
  html_document: default
  word_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#### dataset: Durable Goods
#### author: Dhanshri, Mishitha, Nikhila, Sunanda
### Importing libraries

```{r}
options(warn = -1)
library(data.table)
library(date)
library(dplyr)
library(MASS)
library(devtools)
library(clusplus)
library(plotly)

```

### Reading file and formatting

```{r}
data_transaction = fread("DurData1_Final-2.csv")
names(data_transaction)
summary(data_transaction)
data_transaction$FormattedDate <- as.Date(data_transaction$Transaction_Date, "%d%b%Y:%H:%M:%S")
date0 = as.Date("2004-11-30")
class(date0)
```

### Selecting columns to caluculate RFM

```{r}
data_transaction_test = data_transaction[, c(1,6, 7, 16, 39 )]
data_transaction_test$Recency_Test = date0 - data_transaction_test$FormattedDate
```

### RFM calculation

```{r}
data_transaction_agg = data_transaction_test[ ,.(FrequencyTemp = length(unique(ORIGINAL_TICKET_NBR))), by=.(Household_ID,FormattedDate)]
data_transaction_agg = data_transaction_agg[ ,.(Frequency = sum(FrequencyTemp)), by=Household_ID]
nrow(data_transaction_agg)
data_transaction_agg_test= data_transaction_test[ ,.(Recency= min(Recency_Test),Monetary = sum(EXTENDED_PRICE)), by=Household_ID]
nrow(data_transaction_agg_test)
data_transaction_agg<- merge(data_transaction_agg_test, data_transaction_agg, by=c("Household_ID"), all=TRUE)
nrow(data_transaction_agg)
```

### Removing monetary values which is negative(Removing households where we don’t have purchase transactions for its returns)

```{r}
data_transaction_agg = data_transaction_agg[data_transaction_agg$Monetary>=0,]
```

### Calculating RFM scores 

```{r}
data_transaction_agg$Recency = as.numeric(data_transaction_agg$Recency)
data_transaction_agg$Recency_Score = cut(data_transaction_agg$Recency, breaks = 5 ,labels = 5:1)
data_transaction_agg$Frequency_Score = cut(data_transaction_agg$Frequency, breaks = 5, labels = 1:5)
data_transaction_agg$Monetary_Score = cut(data_transaction_agg$Monetary, breaks = 5, labels = 1:5)
summary(data_transaction_agg)

#converting scores to numeric
data_transaction_agg$Recency_Score = as.numeric(data_transaction_agg$Recency_Score)
data_transaction_agg$Frequency_Score = as.numeric(data_transaction_agg$Frequency_Score)
data_transaction_agg$Monetary_Score = as.numeric(data_transaction_agg$Monetary_Score)
```

### Clustering

```{r}

k = 3
set.seed(100)
cus_km_3 = kmeans(data_transaction_agg[,c(5:7)],k,nstart = 25)
data_transaction_agg[, seg := cus_km_3$cluster]
data_transaction_agg[, .N, seg][order(seg)]
withinss3=sum(cus_km_3$withinss)
rfm_df = names(data_transaction_agg[,c(5:7)] )
data_transaction_agg[, lapply(.SD,mean), seg, .SDcols = rfm_df ][order(seg)]

k = 4
set.seed(100)
cus_km_4 = kmeans(data_transaction_agg[,c(5:7)],k,nstart = 25)
data_transaction_agg[, seg := cus_km_4$cluster]
data_transaction_agg[, .N, seg][order(seg)]
withinss4=sum(cus_km_4$withinss)
rfm_df = names(data_transaction_agg[,c(5:7)] )
data_transaction_agg[, lapply(.SD,mean), seg, .SDcols = rfm_df ][order(seg)]


k = 5
set.seed(100)
cus_km_5 = kmeans(data_transaction_agg[,c(5:7)],k,nstart = 25)
data_transaction_agg[, seg := cus_km_5$cluster]
data_transaction_agg[, .N, seg][order(seg)]
withinss5=sum(cus_km_5$withinss)
rfm_df = names(data_transaction_agg[,c(5:7)] )
data_transaction_agg[, lapply(.SD,mean), seg, .SDcols = rfm_df ][order(seg)]


k = 6
set.seed(100)
cus_km_6 = kmeans(data_transaction_agg[,c(5:7)],k,nstart = 25)
data_transaction_agg[, seg := cus_km_6$cluster]
data_transaction_agg[, .N, seg][order(seg)]
withinss6=sum(cus_km_6$withinss)
rfm_df = names(data_transaction_agg[,c(5:7)] )
data_transaction_agg[, lapply(.SD,mean), seg, .SDcols = rfm_df ][order(seg)]

k = 7
set.seed(100)
cus_km_7 = kmeans(data_transaction_agg[,c(5:7)],k,nstart = 25)
data_transaction_agg[, seg := cus_km_7$cluster]
data_transaction_agg[, .N, seg][order(seg)]
withinss7=sum(cus_km_7$withinss)
rfm_df = names(data_transaction_agg[,c(5:7)] )
data_transaction_agg[, lapply(.SD,mean), seg, .SDcols = rfm_df ][order(seg)]

k = 6
set.seed(100)
cus_km_6 = kmeans(data_transaction_agg[,c(5:7)],k,nstart = 25)
data_transaction_agg[, seg := cus_km_6$cluster]
data_transaction_agg[, .N, seg][order(seg)]
withinss6=sum(cus_km_6$withinss)

#Hartigan rule
hartigan=((withinss3/withinss4)-1)*(data_transaction_agg[,.N]-3-1)
hartigan=((withinss4/withinss5)-1)*(data_transaction_agg[,.N]-4-1)
hartigan=((withinss5/withinss6)-1)*(data_transaction_agg[,.N]-5-1)
hartigan=((withinss6/withinss7)-1)*(data_transaction_agg[,.N]-6-1)

```

### Mean of each segments

```{r}
rfm_df = names(data_transaction_agg[,c(5:7)] )
data_transaction_agg[, lapply(.SD,mean), seg, .SDcols = rfm_df ][order(seg)]

```

### Plot for RFM
```{r}
plot_clus_coord(cus_km_6, data_transaction_agg[,c(5:7)])

```


### aggregation for demo variables
```{r}
data_extrademo_agg = data_transaction[,.(Total_transactions = mean(Transaction_Total), Total_online = sum(Online_Transaction),Total_returns = sum(Quantity<0), Total_diff_categories = length(unique(Category_Description))),by=Household_ID]

#Merging to get demographic from base data
subsetdemo<- data_transaction[,c(19:38)]
subsetdemo<- cbind(data_transaction[,1],subsetdemo)
#Taking unique rows of subsetdemo
demodata<-distinct(subsetdemo)
data_demographic_agg<- merge(demodata, data_extrademo_agg, by=c("Household_ID"), all=TRUE)
nrow(data_demographic_agg)
summary(data_demographic_agg)

#Making gender, no of children
data_demographic_agg$Gender = ifelse(data_demographic_agg$Gender_Individual == "F", 1,0)
data_demographic_agg$No_of_Children = rowSums(data_demographic_agg[,c(7:21)])
names(data_demographic_agg)
lda_data = data_demographic_agg[,c(1,2,4,22:27)]

#Just taking offline transactions to calculate no ofstores visited
offline_transaction_df = subset(data_transaction,Online_Transaction !=1)
data_offline_stores= offline_transaction_df[,.(No_of_stores =length(unique(Transaction_Location)) ),by=Household_ID]
nrow(data_offline_stores)

```

### Creating final dataframe for lda

```{r}
nrow(lda_data)
lda_data_final = merge(x = lda_data, y = data_offline_stores, by = "Household_ID", all.x = TRUE)

#final dataframe with RFM and demo variables.
final_df = merge(x = data_transaction_agg, y = lda_data_final, by = "Household_ID", all.x = TRUE)
nrow(final_df)
names(final_df)

```

### LDA

```{r}
#finding out no of NA
apply(is.na(final_df),2,sum)

#removing NA
nrow(final_df)
final_df_Na_omit=final_df[complete.cases(final_df),]
nrow(final_df_Na_omit)

lda_data_names = names(final_df[,9:17, with = FALSE])
lda_data_names
demo_mean=final_df_Na_omit[, lapply( .SD , mean) , .SDcols = lda_data_names, seg][order(seg)]
print(t(demo_mean))
```

### CLV

```{r}
# Calculating Mean of Recency
Rec_mean = mean(data_transaction_agg[["Recency"]])
```

### Calculate average order value(Total Revenue/no of orders)

```{r}
data_transaction_agg$Avg_order_price = data_transaction_agg$Monetary/data_transaction_agg$Frequency
nrow(data_transaction_agg)
```

### Calculate purchase frequency(Total no of order/No of unique customers )
```{r}
total_purchases = sum(data_transaction_agg$Frequency)
purchase_freq = total_purchases / (length(data_transaction_agg$Household_ID))
purchase_freq
```

### Calculate customer value(Avg order value * purchase frequency)

```{r}
data_transaction_agg$customer_value = data_transaction_agg$Avg_order_price * purchase_freq
```

### Calculate churn column

```{r}
data_transaction_agg$churn = ifelse(data_transaction_agg$Recency > Rec_mean, 0, 1)
```

### Calculate churn rate

```{r}
churn_rate = sum(data_transaction_agg$churn) / length(data_transaction_agg$Household_ID)
churn_rate
```
### Calculate retention rate 

```{r}
retention_rate = 1 - churn_rate
retention_rate
```
### Calculate avg_lifetime_span

```{r}
avg_lifetime_span = 1 / retention_rate
avg_lifetime_span
```

### Calculate clv 

```{r}
data_transaction_agg$CLV = data_transaction_agg$customer_value * avg_lifetime_span

```

### Mean of each segments with CLV

```{r}
rfm_df = names(data_transaction_agg[,c(5:7,12)] )
plot_data=data_transaction_agg[, lapply(.SD,mean), seg, .SDcols = rfm_df ][order(seg)]

```

### Plots for finding
```{r}
#Plot with CLV values
a <- list(
  x = list(plot_data$seg[1], plot_data$seg[4], plot_data$seg[5], plot_data$seg[6]),
  y = list(plot_data$CLV[1], plot_data$CLV[4], plot_data$CLV[5], plot_data$CLV[6]),
  text = list("Online Customer", "Premium Customer", "New Customer", "Valuable Customer"),
  xref = "x",
  yref = "y",
  showarrow = TRUE,
  arrowhead = 7,
  ax = 10,
  ay = -40)

plot_ly(plot_data, x = plot_data$seg, y = plot_data$CLV, type = 'bar', 
        marker = list(color = c('rgb(255,153,0)','rgb(204,0,0)','rgb(204,0,0)','rgb(0,153,51)','rgb(51,51,204)','rgb(255,204,0)'),
                      line = list(color = 'rgb(0,0,102)',
                                  width = 1.5))) %>%

 layout(title = "CLV vs Segments",
         xaxis = list(title = "Segments"),
         yaxis = list(title = "CLV"),
        legend = list(orientation = "h",  
                     xanchor = "center",  
                     x = 0.5),
        annotations=a)


```






  
  
  

  
  
  
  
  
  
  












































  
  
  
  
  
  
  
  
  
  





  
  
  
  
  
  
  
  
  
  
  
































